Binning Strategies and Related Loss for Large Data

### 2015-02-26 Paper.Rnw
standard binning section edits

We \sout{will} first define a binning algorithm that \hh{allocates}  observed values to equal sized bins in a deterministic manner.
We then define a set of bin centers $\{ x_j^\ast \}$, with $1 \le j \le N_X$, where $N_X$ is the overall number of bins in dimension $X$.  To define the set of bin centers we need only specify the center of the first bin, $x^\ast_1$ and the binwidths, $\omega_X$. 
\hh{I am still not sure how bin centers are defined in the end. XXX}

@@

Mathematically, binning corresponds to a function $b_X(.)$ that assigns to a value $x_i$ the closest bin center $x^\ast_j$ for some bin $j \in \{1, ..., N_X\}$, i.e. the binning function   $b_X(.) : x_i \rightarrow x^\ast_j$ is defined as 
% \begin{center}
% $x^\ast_i = b_X(x_i) = x_{(1)}^\ast + \omega_X \cdot \left\lfloor (x_i-x_{(1)})/\omega_x \right\rfloor$ \hspace{.5cm} $\forall x_i < x_{(n)}$, and \\ 
% \vspace{.03}
% $x^\ast_i = b_X(x_i) = x_{(n)} - \omega_x/2$ \hspace{.5cm} $\forall x_i = x_{(n)}$
% \end{center}
\begin{eqnarray}\label{standbin}
b_X(x_i) =  x_{1}^\ast + \omega_X \cdot \left\lfloor (x_i-x_{1}^\ast)/\omega_x + 1/2 \right\rfloor
\end{eqnarray}

where $x_1^\ast$ is the center of the first bin and $\lfloor . \rfloor$ denotes the floor function, $\lfloor x \rfloor$ is the largest integer value smaller than $x$. 
\hh{something went wrong with the next sentence}
An intuitive understanding of the formula is developed by noting $\left\lfloor (x_i-x_{1}^\ast)/\omega_x + 1/2 \right\rfloor$ represents the number of bin boundaries lie between $x_{1}^\ast$ and $x_i$.  Thus, the binning function is finding the location of the bin center for $x_i$ by adding the distance between $x_{1}^\ast$ and the center for $x_i$ to the known first bin center $x_{1}^\ast$.

\hh{get rid of the woulds, shoulds, cans, ...}
A special case of standard binning can be used to divide the data range into $N_X$ equal sized bins.  To do so we use the bin width, $\omega_X = (x_{(n)} - x_{(1)})/N_X$, where $x_{(1)}$ and $x_{(n)}$ denote the observed minimum and the maximum of $X$, respectively. The first bin would then be centered at $x_{(1)} + \omega_X/2$.

>>>>

\hh{I am still not sure how bin centers are defined in the end. XXX}

@@

\ktm{  We define the process that allocates observed values, $x_i$ for $i \in \{1, ..., n\}$, to equal sized bins in a deterministic manner as a standard binning algorithm. The bins are controlled by two parameters: the center of the first bin, $x^\ast_1$ and the binwidth, $\omega_X$. This leads to the set of bin centers $\{ x_j^\ast =  x_1^\ast + (j-1)\omega_X \text{ for } j \in \{1, ..., N_X\} \}$,  where $N_X$ is the overall number of bins.  }

Mathematically, the standard binning algorithm corresponds to a function $b_X(.)$ that assigns to a value $x_i$ the closest bin center $x^\ast_j$ for some bin $j \in \{1, ..., N_X\}$, i.e. the binning function   $b_X(.) : x_i \rightarrow x^\ast_j$ is defined as 
% \begin{center}
% $x^\ast_i = b_X(x_i) = x_{(1)}^\ast + \omega_X \cdot \left\lfloor (x_i-x_{(1)})/\omega_x \right\rfloor$ \hspace{.5cm} $\forall x_i < x_{(n)}$, and \\ 
% \vspace{.03}
% $x^\ast_i = b_X(x_i) = x_{(n)} - \omega_x/2$ \hspace{.5cm} $\forall x_i = x_{(n)}$
% \end{center}
\begin{eqnarray}\label{standbin}
b_X(x_i) =  x_{1}^\ast + \omega_X \cdot \left\lfloor (x_i-x_{1}^\ast)/\omega_x + 1/2 \right\rfloor
\end{eqnarray}

where $x_1^\ast$ is the center of the first bin and $\lfloor . \rfloor$ denotes the floor function, $\lfloor x \rfloor$ is the largest integer value smaller than $x$.  
\ktm{ This is equivalent to treating the $j^{th}$ bin as a half-open interval that contains all observations $x_i \in [x_j^\ast - \omega_X /2 , x_j^\ast + \omega_X /2  ) $. }
%\hh{something went wrong with the next sentence}
%An intuitive understanding of the formula is developed by noting $\left\lfloor (x_i-x_{1}^\ast)/\omega_x + 1/2 \right\rfloor$ represents the number of bin boundaries lie between $x_{1}^\ast$ and $x_i$.  Thus, the binning function is finding the location of the bin center for $x_i$ by adding the distance between $x_{1}^\ast$ and the center for $x_i$ to the known first bin center $x_{1}^\ast$.

\ktm{ A special case of standard binning divides the data range into $N_X$ equal sized bins.  This leads to the bin width, $\omega_X = (x_{(n)} - x_{(1)})/N_X$, where $x_{(1)}$ and $x_{(n)}$ denote the observed minimum and the maximum of $X$, respectively. The first bin would then be centered at $x_{(1)} + \omega_X/2$. }
